Introducing statistics with \texttt{intRo}

### 2014-11-10 student-paper-competition-2014/paper.Rmd
exciting section started

Beyong being simple, \texttt{intRo} is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. In figure~\ref{fig:ui}, we highlight the five elements that comprise the intRo interface.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/ui_annotate.png}
\label{fig:ui}
\caption{The five elements that comprise the intRo application.}
\end{figure}

 - **Top Navigation** - The top navigation bar includes two sets of buttons. The left-aligned buttons are informational buttons. The first is a link to intRo. The second is a link to this documentation page. The third is a link to the GitHub repository where the code for \texttt{intRo} is housed. The final button is a link to our websites, which contain contact information if there are any questions or comments. The right-aligned buttons are \texttt{intRo} utilities. The first is a button to toggle the visibility of the code panel (5). The last is a button to print the active module results, and the associated code (if visible).
 - **Side Navigation** - The side navigation panel includes the modules which are detailed later in the tutorial.
 - **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
 - **Results Panel** - The results pane displays the result of the selected module and options.
 - **Code Panel** - The code panel displays the R code used to generate the results from the student's \texttt{intRo} session. The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.

There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface. Continue with the tutorial to read more about each one.

## Extensible

# In the classroom
\texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course. Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students. Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis. JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows. Using R itself in the classroom is an appealing alternative, but suffers from some other problems. Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis. When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (This experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student.)

We believe that \texttt{intRo} targets these key limitations. The interface is designed to be quick, seamless, and easy to understand. Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device. R underlies the analysis, but users need not be familiar with R or programming to perform it. The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done. For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.

>>>>

Beyong being simple, \texttt{intRo} is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. In figure \ref{fig:ui}, we highlight the five elements that comprise the intRo interface.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
\label{fig:ui}
\caption{The five elements that comprise the intRo application.}
\end{figure}

 - **Top Navigation** - The top navigation bar includes two sets of buttons. The left-aligned buttons are informational buttons. The first is a link to intRo. The second is a link to this documentation page. The third is a link to the GitHub repository where the code for \texttt{intRo} is housed. The final button is a link to our websites, which contain contact information if there are any questions or comments. The right-aligned buttons are \texttt{intRo} utilities. The first is a button to toggle the visibility of the code panel (5). The last is a button to print the active module results, and the associated code (if visible).
 - **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
 - **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
 - **Results Panel** - The results pane displays the result of the selected module and options.
 - **Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session. The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.

There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface. Each module is placed within a higher level category to further help students navigate the tool. The default modules included in \texttt{intRo} are found in table \ref{tab:modules}, and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.

\begin{table}
\centering
\begin{tabular}[ht!]{c|c|p{.6\textwidth}}
Category & Module & Description \\
\hline
\multirow{2}{*}{Data} & Sources & \\
& Transformation & \\
\hline
\multirow{2}{*}{Summaries} & Graphical & \\
& Numerical & \\
\hline
\multirow{3}{*}{Statistics} & Contingency & \\
& Regression & \\
& T test & 
\end{tabular}
\label{tab:modules}
\caption{A list of default modules that ship with \texttt{intRo} as well as brief descriptions of each.}
\end{table}

## Exciting
In general, introduction to statistics is a class undergraduates dread. It's hard, the language is weird, and the software is hard to use. With \texttt{intRo}, we hope to flip this paradigm. \texttt{intRo} is a fun, easy to use tool that's available on the web, and is even usable on mobile devices. As a web-based application, this tool is immediately more familiar to students than a desktop application. With interactive plots from \texttt{ggvis}, \texttt{intRo} brings a `wow'-factor to learning introductory statistics.

Once students get pulled in by an easy to use interface and fun interaction with their data, \texttt{intRo} has an ulterior motive: to get students excited about programming. By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally. This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.

## Extensible
\label{sec:extensible}

# In the classroom
\texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course. Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students. Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis. JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows. Using \texttt{R} itself in the classroom is an appealing alternative, but suffers from some other problems. Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis. When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (This experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student.)

We believe that \texttt{intRo} targets these key limitations. The interface is designed to be quick, seamless, and easy to understand. Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device. \texttt{R} underlies the analysis, but users need not be familiar with \texttt{R} or programming to perform it. The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done. For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.
